New lower bounds for Schur and weak Schur numbers

### 2021-04-08 article/sat.tex
formatage code

\subsection{The search space previously used in computer search for lower bounds may not contain the optimal partitions}

Rowley's new lower bound for \(WS(6)\) (642)  \hyperlink{label2}{\textbf{[2]}} was a quite significant improvement upon the former
best known lower bound (582) \hyperlink{label7}{\textbf{[7]}}. This previous lower bound was found using a computer (often with Monte-Carlo methods) and by making the
assumption that a good partition for \(WS(n+1)\) starts with a good partition for \(WS(n)\) which is true for small values of \(n\).
Therefore, one may wonder whether the limiting factor are the assumptions or the methods used to search for partitions. It appears that the search space
induced by these assumptions does not contain the optimal partitions.

>>>>

\subsection{The search space previously used in computer search for lower bounds may not contain the optimal partitions}
Rowley's new lower bound for \(WS(6)\) (642) \hyperlink{label2}{\textbf{[2]}} was a quite significant improvement upon
the former
best known lower bound (582) \hyperlink{label7}{\textbf{[7]}}. This previous lower bound was found using a computer
(often with Monte-Carlo methods) and by making the
assumption that a good partition for \(WS(n+1)\) starts with a good partition for \(WS(n)\) which is true for small
values of \(n\).
Therefore, one may wonder whether the limiting factor are the assumptions or the methods used to search for partitions.
It appears that the search space
induced by these assumptions does not contain the optimal partitions.

### 2021-04-08 article/schur.tex
formatage code

\begin{definition}
	We call SF-template of \(n\) colors and lenght \(p\) a partition of \( [\![1,p]\!]\) into \(n\) sum-free subsets \(A_1, A_2, ..., A_n\) which verify :
	\[
	\forall i \in [\![1, n-1]\!], \forall (x,y) \in A_i^2, x+y > p
	\Longrightarrow x+y-p \notin A_i
	\]
	We note \(S_+(n)\) the maximal lenght for a SF-template of \(n\) colors. 
\end{definition}

@@

\begin{proposition}
	Let \(n \in [\![2, +\infty]\!]\), we have :
	\[
	2S(n-1)+1 \leqslant S_+(n) \leqslant S(n)
	\]
\end{proposition}

@@

\begin{remark}
	\(S_+\) and \(S\) have the same asymptotic growth rate.
\end{remark}

\subsection{Inequalities using \(S_+\)}

The main result on \(S_+\) follows. It allows us to improve lower bounds on Schur numbers by computing \(S_+\).

\begin{theorem}
	Let $(n,k), (p,q) \in (\mathbb{N}^*)^2$. If there exists a \(SF\)-template of $k+1$ colors and lenght \(p\),
	and a partition of $n$ sum-free subsets of $[\![1,q]\!]$ then there exists a partition of $n+k$ sum-free subsets of $[\![1,pq]\!]$.
\end{theorem}

Setting $p = S_+(k+1)$ and $q = S(n)$ yields the following corollary.

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S(n+k) \geqslant S_+(k+1)S(n) \]
\end{corollary}

Using a SAT solvers, we are able to exhibit SF-template,
hence providing lower bound on \(S_+\). Moreover, when using 
a SF-template in the above formula, it is possible to get an additive constant. We seek templates providing "affine"
inequalities : \(aS(n+k) + b \) where \(a\) is a lower bound for \(S_+(k+1)\). Further details can be found in the \hyperref[SAT]{SAT section}.
Here are the best inequalities on Schur numbers so far :
\[ S(n+1) \geqslant 3S(n) + 1 \]
\[ S(n+2) \geqslant 9S(n) + 4 \]
\[ S(n+3) \geqslant 33S(n) + 6 \]
\[ S(n+4) \geqslant 111S(n) + 43 \]
\[ S(n+5) \geqslant 380S(n) + 148 \]
\[ S(n+6) \geqslant 1140S(n) + 528 \]

@@

We also have a similar theorem where only \(S_+\) is involved.

\begin{theorem}
	Let $(n,k), (p,q) \in (\mathbb{N}^*)^2$. If there exists a \(SF\)-template of $k+1$ colors and lenght \(p\),
	and \(SF\)-template of \(n\) color and lenght \(q\), then there exists \(SF\)-template of \((n+k)\) and lenght \(pq\).
\end{theorem}

@@

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S_+(n+k) \geqslant S_+(k+1)S_+(n) \]
\end{corollary}

>>>>

\begin{definition}
We call SF-template of \(n\) colors and lenght \(p\) a partition of \( [\![1,p]\!]\) into \(n\) sum-free subsets \(A_1,
A_2, ..., A_n\) which verify :
	\[
	\forall i \in [\![1, n-1]\!], \forall (x,y) \in A_i^2, x+y > p
	\Longrightarrow x+y-p \notin A_i
	\]
	We note \(S^+(n)\) the maximal lenght for a SF-template of \(n\) colors.
\end{definition}

@@

\begin{proposition}
	Let \(n \in [\![2, +\infty]\!]\), we have :
	\[
	2S(n-1)+1 \leqslant S^+(n) \leqslant S(n)
	\]
\end{proposition}

@@

\begin{remark}
	\(S^+\) and \(S\) have the same asymptotic growth rate.
\end{remark}

\subsection{Inequalities using \(S^+\)}

The main result on \(S^+\) follows. It allows us to improve lower bounds on Schur numbers by computing \(S^+\).

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template of \(k+1\) colors and lenght \(p\),
and a partition of \(n\) sum-free subsets of \([\![1,q]\!]\) then there exists a partition of \(n+k\) sum-free subsets
of \([\![1,pq]\!]\).
\end{theorem}

Setting \(p = S^+(k+1)\) and \(q = S(n)\) yields the following corollary.

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S(n+k) \geqslant S^+(k+1)S(n) \]
\end{corollary}

Using a SAT solvers, we are able to exhibit SF-template,
hence providing lower bound on \(S^+\). Moreover, when using
a SF-template in the above formula, it is possible to get an additive constant. We seek templates providing "affine"
inequalities : \(aS(n+k) + b \) where \(a\) is a lower bound for \(S^+(k+1)\). Further details can be found in the
\hyperref[SAT]{SAT section}.
Here are the best inequalities on Schur numbers so far :
\[ S(n+1) \geqslant 3S(n) + 1 \]
\[ S(n+2) \geqslant 9S(n) + 4 \]
\[ S(n+3) \geqslant 33S(n) + 6 \]
\[ S(n+4) \geqslant 111S(n) + 43 \]
\[ S(n+5) \geqslant 380S(n) + 148 \]
\[ S(n+6) \geqslant 1140S(n) + 528 \]

@@

We also have a similar theorem where only \(S^+\) is involved.

\begin{theorem}
	Let \((n,k), (p,q) \in (\mathbb{N}^*)^2\). If there exists a \(SF\)-template of \(k+1\) colors and lenght \(p\),
	and \(SF\)-template of \(n\) color and lenght \(q\), then there exists \(SF\)-template of \((n+k)\) and lenght \(pq\).
\end{theorem}

@@

\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\), we have \\
	\[ S^+(n+k) \geqslant S^+(k+1)S^+(n) \]
\end{corollary}

### 2021-04-08 article/weakschur.tex
formatage code

In this section, we generalize Rowley's constructions in [2]. We then introduce, by analogy with the third section, the integer \(WS_+(n)\) 
to build suitable templates.

@@

	Let $(p,q), (n,k) \in (\mathbb{N}^*)^2$. If there exists a partition of $[\![1,q]\!]$ into $n$ weakly sum-free subsets and a partition of $[\![1,p]\!]$ into $k$ sum-free 
	subsets then there exists a partition of $[\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]$ into $n+k$ weakly sum-free subsets.

@@

In particular, if we choose $q = WS(n)$ and $p = S(k)$ in the last theorem, the next corollary follows.

@@

	$ \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } WS(n+k) \geqslant S(k) \left (WS(n) + \left \lceil \frac{WS(n)}{2} \right \rceil +1 \right) + WS(n)$

@@

\textsc{Proof :} Let $(p,q), (n,k) \in (\mathbb{N}^*)^2$,  $N = p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q$, $\alpha = \left \lceil \frac{q}{2} \right \rceil > 0$ and $\beta = q + \alpha + 1$. 
We denote by $f$ the projection of the equivalence relation induced by the partition of $[\![1,q]\!]$ and $g$ the one induced by the partition of $[\![1,p]\!]$. Each equivalence class is represented by a single integer, therefore :

@@

\[g : [\![1,p]\!] \longrightarrow [\![1,k]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } f(x) = f(y) \Longrightarrow f(x+y) \neq f(x)

@@

Let us start by parting the integers of $[\![1,N]\!]$ in two subsets $\mathcal{A}$ and $\mathcal{B}$ where $\mathcal{A} = [\![1,\alpha]\!] \cup \{a\beta + u \mid (a,u) \in [\![0,p]\!] \times [\![\alpha + 1,q]\!]\}$ and $\mathcal{B} = \{a\beta + u \mid (a,u) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]\}$.\\

@@

First, \underline{$\mathcal{A} \cap \mathcal{B} = \varnothing$} : \\
By contradiction, suppose there exists $x \in \mathcal{A} \cap \mathcal{B} \neq \varnothing $. Then there are $(a,u) \in [\![0,p]\!] \times [\![\alpha + 1,q]\!]$ and $(b,v) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]$ such that $x = a\beta + u = b\beta +v$. By definition of $\alpha$ and  $\beta$ we have $u \in [\![\alpha + 1,q]\!] \subset [\![0,\beta - 1]\!]$. 

@@

	\item If $v \in [\![0,\alpha]\!]$ then $v \in [\![0,\beta - 1]\!]$ and $v \neq u$ because $v < \alpha + 1 \leqslant u$
	\item If $v \in [\![-\alpha,-1]\!]$, we note $\tilde{v} = \beta + v$ and thus have $x = (b-1)\beta + \tilde{v}$ with $\tilde{v} \in [\![\beta - \alpha,\beta - 1]\!] \subset [\![0,\beta - 1]\!]$ and $\tilde{v} \neq u$ because $u < q+1 = \beta - \alpha \leqslant \tilde{v}$.

@@

In either cases, we run into a contradiction because of the remainder's uniqueness in the euclidean division of $x$ by $\beta$.\\

@@

Then, we have \underline{$\mathcal{A} \cup \mathcal{B} = [\![1,N]\!]$}:

@@

	\item On the one hand : $1 = \text{min($\mathcal{A}$)} \leqslant \text{max($\mathcal{A}$)} = p\beta + q = N$ and 
	$1 \leqslant \beta - \alpha = \text{min($\mathcal{B}$)} \leqslant \text{max($\mathcal{B}$)} = p\beta + \alpha \leqslant N$, 
	which gives $\mathcal{A} \cup \mathcal{B} \subset [\![1,N]\!]$.
	\item On the other hand, let $x \in  [\![1,N]\!]$. If $x \leqslant \alpha$, we directly have $x \in \mathcal{A}$, let us then suppose that $x > \alpha$ and write $x = a\beta + u$ 
	the euclidean division of $x$ by $\beta$. We have $x \leqslant N$, thus $a \leqslant p$. We distinguish three cases : \\
	- If $u \in [\![0,\alpha]\!]$ then we necessarily have $a \geqslant 1$ because $x > \alpha$, and so $x \in \mathcal{B}$.\\
	- If $u \in [\![\alpha + 1,q]\!]$, then $x \in \mathcal{A}$. \\
	- If $u \in [\![q + 1,\beta - 1]\!]$ then $x = (a+1)\beta - (\beta - u)$ with $-\alpha \leqslant \beta - u \leqslant 0$. 
	Furthermore, $a \leqslant p - 1$, else we would have $x > N$, and so $x \in \mathcal{B}$ \\
	In any case, $x \in \mathcal{A} \cup \mathcal{B}$ and we can thus conclude that $[\![1,N]\!] \subset \mathcal{A} \cup \mathcal{B}$.

@@

This first partition of $[\![1,N]\!]$ will help us to define our final partition by the projection of its equivalence relation. 
We thereby define $h : [\![1,N]\!] \longrightarrow [\![1,n+k]\!]$ as such :\\
- If $x \in \mathcal{A}$ then $h(x) = f(x \text{ mod } \beta)$ (well defined because $x \text{ mod } \beta \in [\![1,N]\!]$)\\
- If $x \in \mathcal{B}$ then $x = a\beta + u$ with a unique $(a,u) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]$ and we define $h(x) = n + g(a)$\\
The fact that $(\mathcal{A}, \mathcal{B})$ is a partition of $[\![1,N]\!]$ ensures that this definition of $h$ is valid. We then have to verify that $h$ induces weakly sum-free subsets.\\

@@

\underline{The classes of equivalence $h(x)$ for $x \in \mathcal{A}$ are weakly sum-free :}

@@

Let $(x,y) \in \mathcal{A}^2$ such that $h(x) = h(y)$, $x \neq y$ and $x + y \leqslant N$

@@

	\item If $(x,y) \in [\![1,\alpha]\!]^2$ :\\
	We have $x + y \leqslant 2\alpha \leqslant q$ and $x + y = 0\beta + x + y$, therefore $x + y \in \mathcal{A}$. 
	Then, by definition : $h(x) = f(x)$, $h(y) = f(y)$ and $h(x+y) = f(x+y)$, which gives us, thanks to the property verified by $f$, that $h(x+y) \neq h(x)$.
	\item If $(x,y) \in [\![1,\alpha]\!] \times ( \mathcal{A} \text{ \textbackslash} \text{ } [\![1,\alpha]\!] )$ :\\
	We write $y = a\beta + u$ with $(a,u) \in [\![0,p]\!] \times [\![\alpha + 1,q]\!]$. Then $x+y = a\beta + x + u = (a+1)\beta + x + u - \beta$, 
	and if $x + u > q$ it follows that $a \leqslant p-1$ since $x+y \leqslant N$, and $-\alpha \leqslant x + u - \beta \leqslant -1$. 
	Therefore $x+y \in \mathcal{B}$ and $h(x+y) \neq h(x) = f(x)$ by definition of h. On the contrary, if $x - u \leqslant n$, 
	then $x+y \in \mathcal{A}$ and $h(x+y) = f(x+u)$ because $x+u$ is actually the remainder of the euclidean division of $x+y$ by $\beta$. 
	Moreover, $h(x) = f(x)$, $x < u$ and, with our initial hypothesis, $h(x) = h(y) = f(u)$. The property verified by $f$ gives us $f(x+u) \neq f(x)$ which can be rewritten as $h(x+y) \neq h(x)$.
	\item If $(x,y) \in ( \mathcal{A} \text{ \textbackslash} \text{ } [\![1,\alpha]\!] ) \times [\![1,\alpha]\!]$ : \\
	This case is handled exactly like the previous one by swaping the roles of $x$ and $y$.
	\item If $(x,y) \in ( \mathcal{A} \text{ \textbackslash} \text{ } [\![1,\alpha]\!] )^2$ : \\
	We write $x = a\beta + u$ and $y = b\beta + v$ with $(a,u)$ and $(b,v)$ in $[\![0,p]\!] \times [\![\alpha + 1,q]\!]$. Then $x+y = (a+b)\beta + u+v = (a+b+1)\beta + u + v - \beta$ 
	with $a+b \leqslant p-1$ (else we would have $x+y > N$ because $u+v > q$) and $-\alpha \leqslant u + v - \beta \leqslant \alpha$, therefore $x+y \in \mathcal{B}$ and by definition $h(x+y) \neq h(x)$.

@@

In any case, $h(x+y) \neq h(x)$ and the classes of equivalence $h(x)$ for $x \in \mathcal{A}$ are weakly sum-free.\\

@@

\underline{The classes of equivalence $h(x)$ for $x \in \mathcal{B}$ are weakly sum-free :}

@@

Let $(x,y) \in \mathcal{B}^2$ such that $h(x) = h(y)$, $x \neq y$ and $x + y \leqslant N$.\\
We write $x = a\beta + u$ and $y = b\beta + v$ with $(a,u)$ and $(b,v)$ in $[\![1,p]\!] \times [\![-\alpha,\alpha]\!]$. 
We have $h(x) = q + g(a)$ and $h(y) = q + g(b)$, therefore $g(a) = g(b)$. We also have $x+y = (a+b)\beta + u +v$.\\
If $u + v \in [\![-\alpha,\alpha]\!]$, then $x+y \in \mathcal{B}$ and $h(x+y) = g(a+b)$, hence we can deduce that $h(x+y) \neq h(x)$ because of the property verified by $g$. On the contrary, if $u+v \notin [\![-\alpha,\alpha]\!]$, then necessarily $x+y \in \mathcal{A}$. Suppose $x+y \in \mathcal{B}$, then $x+y = c\beta + w$ with $(c,w) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]$. Thus, $c\beta + w = (a+b)\beta + u + v$ and $(a+b-c)\beta = w-u-v$. Furthermore $a+b-c \neq 0$, else we would have $u+v = w \in [\![-\alpha,\alpha]\!]$. This finally leads to the following inequality :

@@

which is absurd. We can therefore conclude that $x+y \in \mathcal{A}$ and by definition of $h$, $h(x+y) \neq h(x)$, proving that the classes of equivalence $h(x)$ for $x \in \mathcal{B}$ are weakly sum-free.\\

@@

Finally, we have showed that every classe of equivalence induced by $h$ is weakly sum-free, which ends the proof.

\begin{remark}
This formula includes the results of Rowley \hyperlink{label2}{\textbf{[2]}} as a special case. For \(n>2\), 
this formula does not give new lower bounds but in the same way as we introduced \(S_+\)\hyperref[SE]{(Definition 3.1)}, 
we will define \(WS_+\) and find inequalities between \(WS_+\),\(WS\) and \(S\)
\end{remark}

\subsection{Definition of \(WS_+\)}
\begin{definition}
Let \( (p,k,b) \in (\mathbb{N}^*)^3\), Let \((A_1,....,A_k)\) a partition of  \([\![1, p]\!]\). 
This partition is said to be a b-weakly-sum-free template (b-WSF-template) of \(k\) colors and lenght \(p\) when:
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\) is weakly-sum-free}
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free}
\\\\
\underline{For \(A_k\) (the special subset):} \quad \(\forall (x,y) \in A_k^2,\)
\\
\[if \quad x+y>b+2(p-b),\quad x+y-2(p-b)\notin A_k\]
\\
\underline{For the others subsets:}\quad \(\forall i \in [\![0,k-1]\!], \forall(x,y) \in A_i^2\)
\\
\[
(Id+(p-b)\mathbf{1}_{ [\![0,b]\!]})(x+y ~\text{mod}~ (p-b)) \notin A_i
\]
\end{definition}

@@

\begin{theorem}
Let \((p,k,b) \in (\mathbb{N}^*)^3\), let \( (q,n) \in (\mathbb{N}^*)^2\). If there exists a partition of \(n\)
 sum-free subsets of \([\![1,q]\!]\) and a partition of k subsets \((A_1,....,A_k)\) of \([\![1, p]\!]\) which is a \(b\)-WSF of \(k\) colors and lenght \(p\), 
 then there exists a partition of \([\![1, b+c+q \times p]\!]\) into \((k+n-1)\) weakly sum-free subsets with c=\(min A_p\bigcap [\![b+1, p]\!] -b-1\).
\end{theorem}

>>>>

In this section, we generalize Rowley's constructions in [2]. We then introduce, by analogy with the third section, the
integer \(WS^+(n)\)
to build suitable templates.

@@

Let \((p,q), (n,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \([\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]\) into \(n+k\)
weakly sum-free subsets.

@@

In particular, if we choose \(q = WS(n)\) and \(p = S(k)\) in the last theorem, the next corollary follows.

@@

\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } WS(n+k) \geqslant S(k) \left (WS(n) + \left \lceil \frac{WS(n)}{2}
\right \rceil +1 \right) + WS(n)\)

@@

\textsc{Proof :} Let \((p,q), (n,k) \in (\mathbb{N}^*)^2\), \(N = p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q\),
\(\alpha = \left \lceil \frac{q}{2} \right \rceil > 0\) and \(\beta = q + \alpha + 1\).
We denote by \(f\) the projection of the equivalence relation induced by the partition of \([\![1,q]\!]\) and \(g\) the
one induced by the partition of \([\![1,p]\!]\). Each equivalence class is represented by a single integer, therefore :

@@

\[g : [\![1,p]\!] \longrightarrow [\![1,k]\!] \text{ and } \forall (x,y) \in [\![1,q]\!]^2 \text{, } f(x) = f(y)

@@

Let us start by parting the integers of \([\![1,N]\!]\) in two subsets \(\mathcal{A}\) and \(\mathcal{B}\) where
\(\mathcal{A} = [\![1,\alpha]\!] \cup \{a\beta + u \mid (a,u) \in [\![0,p]\!] \times [\![\alpha + 1,q]\!]\}\) and
\(\mathcal{B} = \{a\beta + u \mid (a,u) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]\}\).\\

@@

First, \underline{\(\mathcal{A} \cap \mathcal{B} = \varnothing\)} : \\
By contradiction, suppose there exists \(x \in \mathcal{A} \cap \mathcal{B} \neq \varnothing \). Then there are \((a,u)
\in [\![0,p]\!] \times [\![\alpha + 1,q]\!]\) and \((b,v) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]\) such that \(x
= a\beta + u = b\beta +v\). By definition of \(\alpha\) and \(\beta\) we have \(u \in [\![\alpha + 1,q]\!] \subset
[\![0,\beta - 1]\!]\).

@@

\item If \(v \in [\![0,\alpha]\!]\) then \(v \in [\![0,\beta - 1]\!]\) and \(v \neq u\) because \(v < \alpha + 1
\leqslant u\)
\item If \(v \in [\![-\alpha,-1]\!]\), we note \(\tilde{v} = \beta + v\) and thus have \(x = (b-1)\beta + \tilde{v}\)
with \(\tilde{v} \in [\![\beta - \alpha,\beta - 1]\!] \subset [\![0,\beta - 1]\!]\) and \(\tilde{v} \neq u\) because \(u
< q+1 = \beta - \alpha \leqslant \tilde{v}\).

@@

In either cases, we run into a contradiction because of the remainder's uniqueness in the euclidean division of \(x\) by
\(\beta\).\\

@@

Then, we have \underline{\(\mathcal{A} \cup \mathcal{B} = [\![1,N]\!]\)}:

@@

	\item On the one hand : \(1 = \text{min}(\mathcal{A}) \leqslant \text{max}(\mathcal{A}) = p\beta + q = N\) and
\(1 \leqslant \beta - \alpha = \text{min}(\mathcal{B}) \leqslant \text{max}(\mathcal{B}) = p\beta + \alpha \leqslant
N\),
	which gives \(\mathcal{A} \cup \mathcal{B} \subset [\![1,N]\!]\).
\item On the other hand, let \(x \in [\![1,N]\!]\). If \(x \leqslant \alpha\), we directly have \(x \in \mathcal{A}\),
let us then suppose that \(x > \alpha\) and write \(x = a\beta + u\)
the euclidean division of \(x\) by \(\beta\). We have \(x \leqslant N\), thus \(a \leqslant p\). We distinguish three
cases : \\
- If \(u \in [\![0,\alpha]\!]\) then we necessarily have \(a \geqslant 1\) because \(x > \alpha\), and so \(x \in
\mathcal{B}\).\\
	- If \(u \in [\![\alpha + 1,q]\!]\), then \(x \in \mathcal{A}\). \\
- If \(u \in [\![q + 1,\beta - 1]\!]\) then \(x = (a+1)\beta - (\beta - u)\) with \(-\alpha \leqslant \beta - u
\leqslant 0\).
	Furthermore, \(a \leqslant p - 1\), else we would have \(x > N\), and so \(x \in \mathcal{B}\) \\
In any case, \(x \in \mathcal{A} \cup \mathcal{B}\) and we can thus conclude that \([\![1,N]\!] \subset \mathcal{A} \cup
\mathcal{B}\).

@@

This first partition of \([\![1,N]\!]\) will help us to define our final partition by the projection of its equivalence
relation.
We thereby define \(h : [\![1,N]\!] \longrightarrow [\![1,n+k]\!]\) as such :\\
- If \(x \in \mathcal{A}\) then \(h(x) = f(x \text{ mod } \beta)\) (well defined because \(x \text{ mod } \beta \in
[\![1,N]\!]\))\\
- If \(x \in \mathcal{B}\) then \(x = a\beta + u\) with a unique \((a,u) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]\)
and we define \(h(x) = n + g(a)\)\\
The fact that \((\mathcal{A}, \mathcal{B})\) is a partition of \([\![1,N]\!]\) ensures that this definition of \(h\) is
valid. We then have to verify that \(h\) induces weakly sum-free subsets.\\

@@

\underline{The classes of equivalence \(h(x)\) for \(x \in \mathcal{A}\) are weakly sum-free :}

@@

Let \((x,y) \in \mathcal{A}^2\) such that \(h(x) = h(y)\), \(x \neq y\) and \(x + y \leqslant N\)

@@

	\item If \((x,y) \in [\![1,\alpha]\!]^2\) :\\
	We have \(x + y \leqslant 2\alpha \leqslant q\) and \(x + y = 0\beta + x + y\), therefore \(x + y \in \mathcal{A}\).
Then, by definition : \(h(x) = f(x)\), \(h(y) = f(y)\) and \(h(x+y) = f(x+y)\), which gives us, thanks to the property
verified by \(f\), that \(h(x+y) \neq h(x)\).
	\item If \((x,y) \in [\![1,\alpha]\!] \times ( \mathcal{A} \text{ \textbackslash} ~ [\![1,\alpha]\!] )\) :\\
We write \(y = a\beta + u\) with \((a,u) \in [\![0,p]\!] \times [\![\alpha + 1,q]\!]\). Then \(x+y = a\beta + x + u =
(a+1)\beta + x + u - \beta\),
and if \(x + u > q\) it follows that \(a \leqslant p-1\) since \(x+y \leqslant N\), and \(-\alpha \leqslant x + u -
\beta \leqslant -1\).
Therefore \(x+y \in \mathcal{B}\) and \(h(x+y) \neq h(x) = f(x)\) by definition of h. On the contrary, if \(x - u
\leqslant n\),
then \(x+y \in \mathcal{A}\) and \(h(x+y) = f(x+u)\) because \(x+u\) is actually the remainder of the euclidean division
of \(x+y\) by \(\beta\).
Moreover, \(h(x) = f(x)\), \(x < u\) and, with our initial hypothesis, \(h(x) = h(y) = f(u)\). The property verified by
\(f\) gives us \(f(x+u) \neq f(x)\) which can be rewritten as \(h(x+y) \neq h(x)\).
	\item If \((x,y) \in ( \mathcal{A} \text{ \textbackslash} ~ [\![1,\alpha]\!] ) \times [\![1,\alpha]\!]\) : \\
	This case is handled exactly like the previous one by swaping the roles of \(x\) and \(y\).
	\item If \((x,y) \in ( \mathcal{A} \text{ \textbackslash} ~ [\![1,\alpha]\!] )^2\) : \\
We write \(x = a\beta + u\) and \(y = b\beta + v\) with \((a,u)\) and \((b,v)\) in \([\![0,p]\!] \times [\![\alpha +
1,q]\!]\). Then \(x+y = (a+b)\beta + u+v = (a+b+1)\beta + u + v - \beta\)
with \(a+b \leqslant p-1\) (else we would have \(x+y > N\) because \(u+v > q\)) and \(-\alpha \leqslant u + v - \beta
\leqslant \alpha\), therefore \(x+y \in \mathcal{B}\) and by definition \(h(x+y) \neq h(x)\).

@@

In any case, \(h(x+y) \neq h(x)\) and the classes of equivalence \(h(x)\) for \(x \in \mathcal{A}\) are weakly
sum-free.\\

@@

\underline{The classes of equivalence \(h(x)\) for \(x \in \mathcal{B}\) are weakly sum-free :}

@@

Let \((x,y) \in \mathcal{B}^2\) such that \(h(x) = h(y)\), \(x \neq y\) and \(x + y \leqslant N\).\\
We write \(x = a\beta + u\) and \(y = b\beta + v\) with \((a,u)\) and \((b,v)\) in \([\![1,p]\!] \times
[\![-\alpha,\alpha]\!]\).
We have \(h(x) = q + g(a)\) and \(h(y) = q + g(b)\), therefore \(g(a) = g(b)\). We also have \(x+y = (a+b)\beta + u
+v\).\\
If \(u + v \in [\![-\alpha,\alpha]\!]\), then \(x+y \in \mathcal{B}\) and \(h(x+y) = g(a+b)\), hence we can deduce that
\(h(x+y) \neq h(x)\) because of the property verified by \(g\). On the contrary, if \(u+v \notin
[\![-\alpha,\alpha]\!]\), then necessarily \(x+y \in \mathcal{A}\). Suppose \(x+y \in \mathcal{B}\), then \(x+y = c\beta
+ w\) with \((c,w) \in [\![1,p]\!] \times [\![-\alpha,\alpha]\!]\). Thus, \(c\beta + w = (a+b)\beta + u + v\) and
\((a+b-c)\beta = w-u-v\). Furthermore \(a+b-c \neq 0\), else we would have \(u+v = w \in [\![-\alpha,\alpha]\!]\). This
finally leads to the following inequality :

@@

which is absurd. We can therefore conclude that \(x+y \in \mathcal{A}\) and by definition of \(h\), \(h(x+y) \neq
h(x)\), proving that the classes of equivalence \(h(x)\) for \(x \in \mathcal{B}\) are weakly sum-free.\\

@@

Finally, we have showed that every classe of equivalence induced by \(h\) is weakly sum-free, which ends the proof.

\begin{remark}
This formula includes the results of Rowley \hyperlink{label2}{\textbf{[2]}} as a special case. For \(n>2\),
this formula does not give new lower bounds but in the same way as we introduced \(S^+\)\hyperref[SE]{(Definition 3.1)},we will define \(WS^+\) and find inequalities between \(WS^+\),\(WS\) and \(S\)
\end{remark}

\subsection{Definition of \(WS^+\)}
\begin{definition}
Let \( (p,k,b) \in (\mathbb{N}^*)^3\), Let \((A_1,....,A_k)\) a partition of  \([\![1, p]\!]\).
This partition is said to be a b-weakly-sum-free template (b-WSF-template) of \(k\) colors and lenght \(p\) when:
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\) is weakly-sum-free}
\\\\
\underline{\(\forall i \in [\![1, k]\!], \quad A_i\backslash [\![1, b]\!]\) is sum-free}
\\\\
\underline{For \(A_k\) (the special subset):} \quad \(\forall (x,y) \in A_k^2,\)
\\
\[if \quad x+y>b+2(p-b),\quad x+y-2(p-b)\notin A_k\]
\\
\underline{For the others subsets:}\quad \(\forall i \in [\![0,k-1]\!], \forall(x,y) \in A_i^2\)
\\
\[
(Id+(p-b)\mathbf{1}_{ [\![0,b]\!]})(x+y ~\text{mod}~ (p-b)) \notin A_i
\]
\end{definition}

@@

\begin{theorem}
Let \((p,k,b) \in (\mathbb{N}^*)^3\), let \( (q,n) \in (\mathbb{N}^*)^2\). If there exists a partition of \(n\)
sum-free subsets of \([\![1,q]\!]\) and a partition of k subsets \((A_1,....,A_k)\) of \([\![1, p]\!]\) which is a
\(b\)-WSF of \(k\) colors and lenght \(p\),
 then there exists a partition of \([\![1, b+q \times p]\!]\) into \((k+n-1)\) weakly sum-free subsets.
\end{theorem}

In particular, if we choose \(q = S(n)\) and \(p = WS^+(k)\) in the last theorem, the next corollary follows.

\begin{corollary}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } let ~ b_{max}=max \{b\in \mathbb{N}^*/WS^+(k+1)=WS_b^+(k+1)\},\)	\[ WS(n+k) \geqslant S(n) WS^+(k+1)+b_{max}\]
\end{corollary}

\begin{remark}
There exists a \(c(b) \geqslant \text{min}(A_{k+1} \backslash [\![1,b] - b - 1\) such that \(c(b)\) numbers can be
added at the end of the extended partition.Therefore, we can get a better lower bound by finding a couple \((b,c(b))\)
which maximizes the sum \(b+c(b)\) such that \(WS^+(k+1)=WS_b^+(k+1)\). Hence we have:
	\[ WS(n+k) \geqslant S(n) WS^+(k+1)+b+c(b)\]

\end{remark}
